Assessing \hydrogen{} and \carbon{} Network Strategies in Europe

### 2023-12-05 report/submissions/journal/paper/paper.tex
rewrite 3.1 without regional context

\documentclass[twocolumn]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{cuted}  % Add the cuted package
\usepackage{caption}
\captionsetup{font=small}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{fancyhdr,lipsum}
\usepackage{subcaption}

@@

Total annual system costs in the Baseline model are highest, totalling \label{}803 billion euros. The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reductions (\label{}781 billion euros) for the \carbon{} network, and around \label{}4\% cost reductions for the \hydrogen{} network and the hybrid scenario (\label{}765 and \label{}764 billion euros, respectively).

In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the Baseline model), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass. Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros for the Baseline model), with the electricity grid making up more of the system costs than the other three networks combined in all models. Capital expenditures on carbonaceous fuel production are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros). Heat pump installations make up around \label{}8\% (60 billion euros).

@@

\subsection*{Transport systems unlock point-source carbon capture}

@@

We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
CU is preferably deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value). In the \modBase{}, this means combining the lowest-cost CC source (process emissions) with electrolysis in all regions, even if the renewable resources are poor and electrolysis costs are thus relatively high (e.g., non-coastal regions in Germany). In regions with medium-quality renewables (e.g., France and Scandinavia), CC biomass is also deployed -- and combined with medium-price electrolytic hydrogen -- to run additional CU. But the largest volumes of CU are deployed in the regions with the best renewables and lowest-cost electrolysis (e.g., Iberian Peninsula and British Isles), where all point-source CC potentials are exhausted and additionally necessary carbon is sourced from high-cost DAC.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_hydrogen.png}
    \caption{Balance of captured hydrogen across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
    \label{fig:balance_captured_carbon}
\end{figure}

>>>>

\documentclass[twocolumn]{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{cuted}  % Add the cuted package
\usepackage{caption}
\captionsetup{font=small}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tabularx, multirow}
\usepackage{fancyhdr,lipsum}
\usepackage{subcaption}
\usepackage[shortcuts]{extdash}

@@

Total annual system costs in the \modBase{} are highest, totalling \label{}803 billion euros. The models with additional transport options for carbon and hydrogen achieve a more efficient system allocation and lower total system costs, with around \label{}2\% cost reduction (\label{}781 billion euros) in the \modCO{}, and around \label{}4\% cost reduction in the \modH{} and the \modHybrid{} (\label{}765 and \label{}764 billion euros, respectively).

In all models, almost half of the system costs are spent on primary electricity production from wind and solar power (\label{}45\% or \label{}360 of \label{}799 billion euros total costs in the Baseline model), with another \label{}12\% (\label{}100 billion euros) spent on hydro, nuclear and biomass. Costs for energy transport networks for electricity, gas, hydrogen and carbon are below \label{}12\% (\label{}100 billion euros), with the electricity grid making up more of the system costs than the other three networks combined. This is the case for all four models. Capital expenditures on carbonaceous fuel production are around \label{}3\% (\label{}25 billion euros) in each model, while electrolysis makes up around \label{}5\% (\label{}40 billion euros). Heat pump installations account for around \label{}8\% (60 billion euros) of total system costs.

@@

\subsection*{Transport Systems Unlock Point\=/Source Carbon Capture}

@@

% 1. baseline model is sub-optimal -> un-used landlocked carbon capture potentials; scarcity in hydrogen and power in Central Europe due to industrial clusters and high population density; costly DAC + additional renewables at sequestration sites
We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models. In comparison, the \modBase{} reveals significant inefficiencies: On the one hand, regions with no access to sequestration sites (i.e., ``landlocked'' regions) underutilize CC potentials from process emissions, fossil gas and biomass. On the other hand, sequestration sites rely on costly DAC to capture the carbon for CS from the air. Furthermore, hydrogen and power prices show high regional differences, indicating system inefficiencies due to transport constraints caused by non-existent (in case of hydrogen) or insufficient (in case of power) network connections. Prices are significantly higher in Central Europe, a region with high population density and high concentration of energy-intensive industries.

% 2. introduce carbon balance figure -> transport systems unlock point-source CC potentials, in particular biomass and gas for industry, less methanation, less DAC and therefore less DAC specific renewables
Transport systems emerge as a crucial component in optimizing CC. By enabling the efficient transportation of carbon, hydrogen, or both, these models reduce the reliance on high-cost DAC, and source more carbon from low-cost point-source CC on biomass and gas in industrial applications (see Fig.~\ref{fig:balance_captured_carbon}). The \modH{} only captures \label{}200 Mt of \carbon{} from DAC compared to \label{}400 Mt in the \modBase{}. Instead, the \modH{} captures more carbon from process emissions as well as gas and biomass for industry. The \modCO{} uses only \label{}150 Mt from DAC, and instead also captures more carbon than the \modH{} from gas for industry and biomass CHPs. And finally, the \modHybrid{} captures only \label{}100 Mt from DAC, but slightly more carbon from all other CC technologies.
Also, the networks reduce the reliance on methanation in the model, reducing the overall demand for CC volumes. Another knock-on effect of the reduced reliance on DAC is reduced endogenous demand for renewables to power DAC facilities.

% 3. introduce capture share figure -> correlation between average cf and cc share, since gas CHPs and PPs are peak load techs, no cc infrastructure;
The increase in captured carbon from point sources in the network models is mostly due to CC infrastructure being deployed on industrial or power plant facilities that are present in all models (as opposed to the deployment of additional electricity generating capacity for the purpose of carbon capture). This is evident when analyzing the capture share of the different CC technologies (see Fig.~\ref{fig:captureshare_line}).

In the \modBase{}, the capture share is highest for biogas-to-gas facilities, followed by process emissions, biomass for industry, biomass CHPs, gas for industry, SMR and gas CHPs. For the network models, we can observe a correlation between CC share and capacity factors of the underlying technologies: Process emissions as well as gas and biomass for industry have high capacity factors (above 80\%), and CC shares close to 100\% (the only exception being gas for industry in the \modH{}). Biomass CHPs have a capacity factor of around 40\%, and CC shares of 50\% in the \modH{} and 95\% in the \modCO{}. SMR and gas CHPs have capacity factors below 20\%, and CC shares around 50\% and 0\%, respectively, across all network models. Gas CHPs serve as peak load electricity production technology, and with very few operating hours, the high investment costs into CC applications are not efficient.\footnote[1]{Similarly, SMR serves as ``peak-load'' hydrogen production technology in regions with poor renewable resources for electrolysis.}

% 4. carbon-grid has higher absolute CC and CC shares than hydrogen-grid and less DAC however hydrogen-grid is more cost-effective

[old text]
We can now closer examine the carbon capture activity when shifting from the \modBase{} to the other three network models.
CU is preferably deployed in the regions where carbon, hydrogen as well as the inputs and by-product outputs (e.g., waste heat) of the synthesis processes have the lowest cost (or highest value). In the \modBase{}, this means combining the lowest-cost CC source (process emissions) with electrolysis in all regions, even if the renewable resources are poor and electrolysis costs are thus relatively high (e.g., non-coastal regions in Germany). In regions with medium-quality renewables (e.g., France and Scandinavia), CC biomass is also deployed -- and combined with medium-price electrolytic hydrogen -- to run additional CU. But the largest volumes of CU are deployed in the regions with the best renewables and lowest-cost electrolysis (e.g., Iberian Peninsula and British Isles), where all point-source CC potentials are exhausted and additionally necessary carbon is sourced from high-cost DAC.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/balance_bar_hydrogen.png}
    \caption{Balance of captured hydrogen across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
    \label{fig:balance_hydrogen}
\end{figure}
